### 2016-09-13 10_conc.tex

\section{Conclusions and Future Work}
%\section{Discussion and Conclusions}
\label{sec-conc}

% \sm{to be revised according to new material. Also, I guess we should add some more ``meat'' to this section}
% 
% \aht{The failure anaysis of 23 docs with max/min ratio $>$10000 suggests that we might need a 0 in the scale,
% and that we might want a quality check that looks at precision of ratios within a unit. For example if they
% say, 1, 10, 37134: "Can you really be sure that this document is precisely 3.7134 times more relevant than the
% previous?" It also suggests that the normalisation might need to be more agressive.
% }
%

Relevance is a fundamental concept in information retrieval,
underpinning test collection-based system effectiveness evaluation.
This is the first study to have collected large scale, real user data
about relevance perceptions using magnitude estimation, obtaining 
%
through crowdsourcing
%
over 50,000 ratings across 18 TREC topics. 
We close the paper by briefly summarizing our findings and sketching
future work, as well as highlighting some limitations of this study.

\subsection{Summary}
\label{sec:summary}

The first research question that was considered asked whether magnitude
estimation is a suitable technique for judging relevance at the
document level.
A comparison of the overall distribution of magnitude estimation
judgments shows a close alignment with ordinal graded relevance
categories, both at the individual topic level, and across the set of
18 topics as a whole.
A high level of consistency was also established between the
distribution of magnitude estimation judgments and classical binary
relevance assessments.
A failure analysis demonstrated that for those cases where the
magnitude estimation and ordinal judgments disagree, this is typically
not due to problems with magnitude estimation per se, but rather arises
due to different judging context, or from ambiguity in the original
topic statements.

Magnitude estimation is a technique that may be unfamiliar to many
people.
The second research question therefore focused on the suitability of
using crowdsourcing to obtain magnitude estimation judgments at scale.
Aiming to avoid spam workers, our crowd task included four quality
checks: a practice line length estimation task; a multiple choice topic
understanding question; a check that magnitude estimates that were
entered for two documents with known ordinal relevance were rank
consistent; and, a minimum time of 20 seconds spent per document.
The magnitude estimation judgments that were obtained after this
process showed a high level of external agreement with existing TREC
binary and Sormunen ordinal judgments: rank ordered magnitude
estimation scores agree with binary judgments 86\% of the time, and
with ordinal judgments 80\% of the time (the agreement with ordinal
judgments is higher for ordinal levels that are further apart, and
lower for ordinal levels that are closer together).
The ranked agreement between binary and ordinal judgments was 85\%,
giving confidence that the using magnitude estimation is robust as a
scaling technique for document-level relevance.

Considering the impact of magnitude estimation relevance judgments on
IR system evaluation, the third research question, our analysis showed
that results can vary substantially when different relevance scales are
used, with a $\tau$ of 0.677 between system orderings when
effectiveness is measured using nDCG@10 and comparing magnitude
estimation scores with binary relevance across TREC runs.
When comparing magnitude estimation with ordinal relevance on simulated
runs, the correlation is even lower at 0.147.
Moreover, when individual magnitude estimation scores are considered,
rather than the mean, the correlations show a wide range of individual
variation in terms of the perception of relevance.
A key factor appears to be that different searchers have different
perceptions of the extent of relevance variation, using wide or thin
scales.
Overall, the analysis suggests that it is important to incorporate
different judging scales, rather that assuming that a single scale can
closely reflect a larger population of users.

We also employed magnitude estimation to directly investigate gain
profiles, the fourth research question.
Typically, gains in nDCG are set as linear or exponential profiles.
Our analysis of user-reported relevance perception showed that the
linear profile is a closer fit to the relevance perceptions of the
``average'' user.
However, the distribution of magnitudes again suggests that attempting
to fit a single profile (or view of relevance) for system evaluation is
unlikely to be sufficient, if the expectation is that the evaluation
should be reflective of a broad user base.

While on average our magnitudes were broadly equivalent to previous
ordinal scales, the outstanding feature of our data was the wide range
of scores that participants chose to employ in the judging task.
In particular, at least half of the participants chose gain values that
are not consistent with currently used values.
Section~\ref{sec:ranking-trec-runs} shows that using judgments made on
a wide scale leads to different system rankings than judgments
collected on a narrow scale.
Recall that these scales are not imposed on the judge, as they are in
all previous relevance judgment tasks in the literature, but are chosen
by the participants themselves.

This is another key contribution of this study: when a priori
categorical scales are used for relevance judgment tasks, there is no
possible way to capture variance in human perception of the scale of
relevance.
In turn, this limits our understanding of how gain should be set in
DCG-like metrics, and hence our ability to accurately evaluate systems.

\subsection{Limitations and Future Work}
\label{sec:sig-figs}

This study is the first to investigate document level relevance
judgments using magnitude estimation, at scale.
A much-cited benefit of magnitude estimation is that it is a ratio
scaling technique.
Indeed, workers were instructed to assign relevance scores as ratios
to those assigned to previous documents.
However, some documents in the collected data lead to ratios that have
a large number of significant figures.
For example, in one unit, the first document was given a score of 54,
and the second 76, implying that the worker thought that the second
document was 1.407 times more relevant than the first.
This seems like an unusually high amount of precision for such a task,
and it may be that the worker was choosing numbers without closely
following the ratio requirement.
Intuitively, one might think that it would be possible to distinguish
the relevance of two documents at a ratio of one significant figure
(for example, ``2 times as relevant'', or ``100 times as relevant'');
or maybe even two significant figures (``21 times as relevant'', or
''140 times as relevant''); but it seems very unlikely that this could
be done to three significant figures or more without extensive training
and effort.
We can safely assume that our crowdsourced workers were not trained in
relevance assessment, nor spent more than a minute or two on each
document (mean time to assess 78 seconds, standard deviation 86
seconds).

If we assess each unit of work using a criterion that each relevance
score must have a ratio with at least one previously assigned score
that has at most two significant figures, then 4806 of the
7059 units match.
If we are more strict, and insist on only one significant figure, then
only 3591 of 7059 units match.
It seems, therefore, that about half of the units are being completed
without a strict adherence to the idea of assigning numbers following 
ratio relationships.
We note, however, that this might not be a problem in practice because of
approximation and balancing effects: the 1.407 figure above is likely
not exact, but may be a good approximation of a value between 1 and 2.
Moreover, relevance is not a strictly defined concept, and the magnitude
estimation is intended to reflect human perceptions.
Overall, the obtained magnitude estimation scores
were shown to have high external consistency with other judging scales.

Throughout the paper we have assumed that topical relevance collected
using magnitude estimation can be used directly as gain in DCG-like
measures.
While perhaps being more representative of user perceptions than other
relevance scales, this still makes many assumptions about the user's
search process which are probably untrue: for example, ignoring the
interdependence of documents and other aspects of relevance.
In future work, we plan to investigate whether magnitude estimation can
be used to reliably scale other aspects of relevance such as novelty,
as well as search outcome measures such as satisfaction, and to examine
whether this can lead to more meaningful gain representations.

In this paper we focused on quantitative analysis of the magnitude
estimation process. As part of the judging interface, judges were also
asked to enter short textual comments to justify their choice of score.
We plan to analyze this qualitative data in future work.

The bulk of the analysis in this paper considered magnitude estimation
judgments as a reflection of relevance at the aggregate level. 
A further interesting line of analysis would be to consider the
technique at the topic level, as different perceptions of relevance may
be useful of indications of topic difficulty, or as potential
calibration functions of how well a system could be expected to perform.

% Local Variables:
% TeX-master: "ME-TOIS.tex"
% End: