### 2016-09-13 1_intro.tex

\section{Introduction}
\label{sec-intro}

Relevance is an important concept in information retrieval (IR), and
relevance judgments form the backbone of test collections, the most widely-used
approach
for the evaluation of IR system effectiveness.
Document relevance judgments are typically made using ordinal scales,
historically at a binary level, and more recently with multiple
levels~\citep{VooHar05}.
However, despite its importance, operationalizing relevance for
the evaluation of IR systems remains a 
complicated issue; for example,
when using an ordinal scale 
 it is unclear how many relevance categories should be
chosen~\cite{TanSha99}.

Magnitude estimation is a psychophysical technique for measuring the
sensation of a stimulus.
Observers  assign numbers to a series of stimuli, such
that the numbers reflect the perceived difference in intensity of each
item.
The outcome is a ratio scale of the subjective perception of the
stimulus; if a magnitude of 50 is assigned to one stimulus, and 10 to
another, then it can be concluded that the two items are perceived in a
5:1 ratio of sensation~\cite{moskowitz:1977}.
While initially developed for the measurement of the sensation of
physical stimuli such as the intensity of a light source, magnitude
estimation has been successfully applied to the measurement of
non-physical stimuli including the usability of
interfaces~\cite{McG03}.

Being able to derive ratio scales of subjective perceptions of the
intensity of stimuli, magnitude estimation may be a useful tool for
measuring and better understanding document relevance judgments.
The application of magnitude estimation in the IR field has been
limited to the consideration of the relevance of carefully curated
abstracts returned from bibliographic databases~\citep{Eis88}, and our
own small-scale pilot study~\cite{SchMad14,MadMiz15}.
While suggestive, these studies have been limited in terms of
demonstrating the broader utility of the approach, or its direct
application to IR evaluation.
In this work, we investigate the larger-scale application of magnitude
estimation to document relevance scaling, reporting on a user study
over 18 TREC topics and obtaining judgments for 4,269 documents.
This is also the first work to consider the direct application of
magnitude estimation to the evaluation of IR systems, and to rely on
crowdsourcing workers to gather a large amount of magnitude estimation
relevance assessments.
Specifically, we consider four research questions.

\begin{enumerate}[RQ1.]
\itemsep 0em

\item \label{item:rq1} Is the magnitude estimation technique suitable
  for gathering document-level relevance judgments, and are the
  resulting relevance scales reasonable with respect to our current
  knowledge of relevance judgments?

\item \label{item:rq4} 
  Is crowdsourcing a viable method to gather robust magnitude
  estimation values?
  Crowdsourcing might exacerbate some of the
  potential complications of using magnitude estimation, such as the
  subjective scale used by each judge, or the difficulty in using a
  ratio scale without proper training.     
  Do the workers agree with
  previous judgments, and are 
  redundant workers required to obtain
  agreement with previous judging methods?

\item \label{item:rq2} How does IR system evaluation change when ratio
  scale magnitude estimation relevance judgments are used to calibrate
  the gain levels of the widely-used nDCG and ERR evaluation metrics,
  compared to using arbitrarily set gain values for a pre-chosen
  number of ordinal levels?

\item \label{item:rq3} Can magnitude estimation relevance scores
  provide additional insight into user perceptions of relevance, into
  actual gain values, and into individual gain profiles?

\end{enumerate}

We also investigate the qualitative comments that users of the magnitude
estimation relevance judging approach were required to make to justify
their scores, and present an analysis of the relationship between
comments, scores and source documents.

In the next section, background material and related work are
presented.
Details of our user study and experimental methodology are provided in
Section~\ref{sec-methods}, and the first descriptive results are
presented in Section~\ref{sec:descr-stat}.
The full analysis and discussion of our results, corresponding to the
four research questions, are detailed in Sections~\ref{sec-rq1}
to~\ref{sec-rq3}.
% Section~\ref{sec:comment-analysis} details the analysis of the comments that
% judges made to justify their magnitude estimation scores.
Our conclusions and directions for future work are included in the
final section of the paper.

% Local Variables:
% TeX-master: "ME-TOIS.tex"
% End: