### 2016-09-13 3_methods.tex

\section{Experimental Methodology}
\label{sec-methods}

In this section we describe the details of the experiments that were
carried out to crowdsource
%gather
the required relevance judgments using magnitude estimation.

\subsection{Topics and Documents}

Document-level relevance assessments have traditionally been made using
binary or multi-level ordinal scales.
To compare magnitude estimation judgments to these approaches, we
selected a set of search tasks and documents from the TREC-8 ad hoc
collection, which studies informational search over newswire documents.
The TREC-8 collection includes binary relevance judgments made by
NIST assessors~\cite{VooHar99}. Subsequently, \citet{Sormunen:2002} carried out a
re-judging exercise where a subset of topics and documents were judged
by a group of six Master's students of information studies, fluent in English although not native speakers,
on a 4-level ordinal relevance scale: \nn--not relevant (0);  
\mm--marginally relevant (1); \rr--relevant (2); \hh--highly relevant
(3).

The 18 \emph{search topics} used in our study were the subset of TREC-8 topics
which also have Sormunen ordinal judgments available. 
They are listed in the first column of Table~\ref{tab:descr-stat}.
The \emph{documents} for which we collected magnitude estimation
judgments were the set of the top-10 items returned by systems that
participated in TREC-8.  
This gives a total of 4,269 topic-document pairs, of
which 3881 have binary TREC relevance judgments available, and 805 
of which have Sormunen ordinal judgments available.
The number of documents for each topic is shown in 
Table~\ref{tab:descr-stat} (2nd column).

\begin{table}[t]%
%   \centering
% %\begin{footnotesize}
% \begin{scriptsize}
\tbl{Topics, number of documents, number of units, \hkh, \nkn, number of back buttons
  usage, number of failures in the (c) and (d) checks, as detailed in
  the text.\label{tab:descr-stat}%
}{%
%  \begin{tabular}{@{}l @{ } r @{ } r  l@{ }l  @{ }  r@{ }r@{ }r  @{ } c@{  } r@{ }r@{ }r@{}}
  \begin{tabular}{l rr  ll rrrcrrr}
\toprule
    Topic&
\multicolumn{1}{l}{No.}  & 
\multicolumn{1}{l}{No.}  &   
\hkh &
\nkn & 
\multicolumn{3}{c}{Back} && 
\multicolumn{3}{c}{Fail}\\ 
    & 
\multicolumn{1}{l}{docs} & 
\multicolumn{1}{l}{units} &    
&  
& 
\multicolumn{1}{c}{$0$} & 
\multicolumn{1}{c}{$1$}& 
\multicolumn{1}{c}{$\ge 2$} && 
\multicolumn{1}{c}{$0$}& 
\multicolumn{1}{c}{$1$}& 
\multicolumn{1}{c}{$\ge2$}\\

% \multicolumn{1}{@{}l@{}}{No.}  & 
% \multicolumn{1}{@{}l@{}}{No.}  &   
% \hkh &
% \nkn & 
% \multicolumn{3}{c}{Back} && 
% \multicolumn{3}{c}{Fail}\\ 
%     & 
% \multicolumn{1}{@{}l@{}}{docs} & 
% \multicolumn{1}{@{}l@{}}{units} &    
% &  
% & 
% \multicolumn{1}{@{}c@{}}{$0$} & 
% \multicolumn{1}{@{}c@{}}{$1$}& 
% \multicolumn{1}{@{}c@{}}{$\ge 2$} && 
% \multicolumn{1}{@{}c@{}}{$0$}& 
% \multicolumn{1}{@{}c@{}}{$1$}& 
% \multicolumn{1}{@{}c@{}}{$\ge2$}\\

\midrule
402	& 278	& 460	& LA111689-0162	& FBIS3-10954       & 452 &  8 & 0 && 426 & 18 & 16\\
403	& 111	& 182	& LA092890-0067	& LA071290-0133     & 178 &  4 & 0 && 120 & 19 & 43\\
405	& 214	& 354	& LA061490-0072	& FBIS3-13680       & 347 &  6 & 1 && 322 & 20 & 12\\
407	& 212	& 350	& FT921-6003	& FR940407-2-00084  & 346 &  3 & 1 && 324 &  9 & 17\\
408	& 188	& 310	& FT923-6110	& LA062290-0070     & 306 &  4 & 0 && 266 & 16 & 28\\
410	& 212	& 350	& FBIS4-64577	& FBIS4-44440       & 346 &  4 & 0 && 326 & 10 & 14\\
415	& 179	& 295	& FBIS3-60025	& FBIS4-10862       & 289 &  4 & 2 && 255 & 23 & 17\\
416	& 174	& 287	& FBIS4-49091	& LA112590-0107     & 286 &  1 & 0 && 267 & 12 &  8\\
418	& 243	& 402	& LA102189-0167	& FT924-6324        & 394 &  8 & 0 && 372 & 15 & 15\\
420	& 164	& 270	& LA121590-0108	& LA112690-0001     & 266 &  4 & 0 && 256 &  4 & 10\\
421	& 342	& 567	& FT941-428	& LA073189-0033     & 554 & 11 & 2 && 537 & 16 & 14\\
427	& 195	& 322	& FT943-5736	& LA080590-0077     & 315 &  7 & 0 && 214 & 52 & 56\\
428	& 253	& 419	& FT943-9226	& FBIS3-20994       & 412 &  6 & 1 && 396 & 13 & 10\\
431	& 203	& 335	& FBIS3-46247	& FT944-5962        & 333 &  2 & 0 && 296 & 27 & 12\\
440	& 264	& 437	& FT942-3471	& LA020589-0074     & 427 & 10 & 0 && 397 & 19 & 21\\
442	& 408	& 677	& LA011390-0057	& FT923-4524        & 672 &  5 & 0 && 629 & 25 & 23\\
445 	& 210	& 347	& FT924-8156	& LA031989-0092     & 334 & 11 & 2 && 334 & 10 & 3 \\
448	& 419	& 695	& LA080190-0139	& FBIS3-16837       & 687 &  8 & 0 && 652 & 21 & 22\\
\addlinespace                                                              
Sum:	&4269	&7059	&		&	            &6944& 106 & 9 &&6389 &329 &341 \\ 
\multicolumn{2}{@{}l@{}}{Percentage:}&&  &                   &  98 &  2 & 0 &&  90 & 5  & 5 \\
\bottomrule
  \end{tabular}%
}%
%\end{scriptsize}
\end{table}

\subsection{User Study}

We carried out a user study through the CrowdFlower crowdsourcing
platform during December 2014 and January 2015.
The experimental
design was reviewed and approved by the RMIT University 
ethics
review board.
Participants were paid \$0.2 for each task \emph{unit}, defined as a group of
magnitude estimation judgments for 8 documents in relation to one
topic. 
The number of units for each topic is shown in
Table~\ref{tab:descr-stat} (3rd column).

\subsubsection{Practice task}
\label{sec:practice-task}

%\myparagraph{Practice task:}
After agreeing to take part in the study, a participant was shown 
a first set of 
task
instructions, including a brief
introduction to the experiment and explanation of the
magnitude estimation process.
Since magnitude estimation may not be familiar to participants, they
were first asked to complete a practice task, making magnitude
estimations of three lines of different lengths, shown one at a time.
If they successfully completed this practice task (success was defined
as assigning magnitudes such that the numbers were in ascending order
when the lines were sorted from shortest to longest), participants were
able to move on to the main part of the experiment.
The full instructions shown to participants for the practice task are
included in Appendix A.

\subsubsection{Main task}
\label{sec:main-task}

%\myparagraph{Main task:} 
The main task of the experiment required
making magnitude estimations of the relevance of documents.
Participants were informed, by means of a second set of instructions,
that they would be shown an information need statement, and then a
sequence of eight documents that had been returned by a search system
in response to the information need statement, presented in an
arbitrary order, and that their task was to 
\emph{indicate how relevant these documents appear} in relation to the information need.

For the main task, the title, description and narrative of a TREC topic
were first displayed at the top of the screen.
After reading the information need, participants had to respond to a
4-way multiple-choice question, to test their understanding of the
topic.
The test questions focused on the main information concepts presented
in the topic statements, and were intended to check that participants
were engaging with the task, and as a mechanism to remove spammers.
Participants who were unable to answer the test question correctly were
unable to continue with the task.

Next, participants were presented with eight documents, one at a time.
For each document, the participant was required to enter a magnitude estimation 
number in a text box displayed directly below the document, and 
a brief justification of why they
entered their number into a larger text field.
They were then able to proceed to the next document.
A back button was available in the interface, with participants being
advised that this should only be used if they wish to correct a mistake;
under 3\% of submitted jobs included use of this feature 
(details on the occurrences of clicks on the back button are shown in
Table~\ref{tab:descr-stat}, where the number of units with zero, one, or two
or more back button clicks are shown).

After entering responses for eight documents, the task was complete.
Participants were able to complete further tasks for other topics, up
to a maximum of 18 tasks, as they were not able to re-assess the same
topic.

\subsubsection{Magnitude estimation assignments}
\label{sec:magn-estim-assignm}

%\myparagraph{Magnitude estimation assignments:} 
Participants were instructed 
to assign ME scores as follows.
\emph{You may use any numbers that seem appropriate to you -- whole numbers,
fractions, or decimals.
However, you may not use negative numbers, or zero.
Don't worry about running out of numbers -- there will always be a
larger number than the largest you use, and a smaller number than the
smallest you use.
Try to judge each document in relation to the previous one.
For example, if the current document seems half as relevant as the
previous one, then assign a score that is half of your previously
assigned score.} 
While some applications of
magnitude estimation use a fixed modulus (specifying a particular number that
is to be assigned to the first stimulus that is presented), this has
been found to promote clustering of responses and the potential
over-representation of some numbers~\cite{moskowitz:1977}; we therefore
allowed participants to freely choose their own values.
The complete
instructions that were shown to participants for both the practice and
main tasks are included in Appendix A.

\subsubsection{Document ordering}
\label{sec:document-ordering}

% \myparagraph{Document ordering:} 
As explained above, each participant
task \emph{unit} required the judging of a set of eight documents for a particular search
topic statement.
The experiments used a randomized design, with documents presented in
random order, to avoid potential ordering effects and first sample
bias~\cite{moskowitz:1977}.
The document sets were constructed such that two of the documents in
each set were a known ordinal \hh and \nn document for the topic;
%
these documents %(referred to in the following as
(henceforth \hkh and \nkn) were the same for all 
the participants working on the same topic 
(the TREC document identifiers for each
topic are shown in Table~\ref{tab:descr-stat}).
This was to ensure that each participant saw at least one document from
the high and low ends of the relevance continuum.
Ensuring that stimuli of different intensity levels are included in a
task has been found to be important for the magnitude estimation
process~\cite{Ges97}, and also has an impact on the score normalization
process, described below.
Moreover, including the two \nkn and \hkh documents with ``known'' ordinal relevance
values enabled a further data collection quality control check: after
judging all eight documents, participants who had assigned magnitude
estimation scores for the \nkn document that were larger than for the
\hkh document were not able to complete the task.
In total, at least 10 ME scores were gathered for each topic-document
pair.

\subsubsection{Quality checks}
\label{sec:quality-checks}

% \myparagraph{Quality checks:} 
In total, four quality checks were included in the
data gathering phase of our experiments: 
%\vspace{-0.5em}
\begin{enumerate}[(a)]
%\itemsep 0em
\item a practice task requiring magnitude estimation of line lengths; 
\item a multiple-choice test question to check a participant's
understanding of the topic; 
\item a check that the magnitude estimations score for \hkh was greater than that
assigned to \nkn; and 
%\end{enumerate}
%\vspace{-0.5em}
%In addition, a time-based quality check was also used: 
%\vspace{-0.5em}
%\begin{enumerate}[(d)]
\item each participant had to spend at least 20 seconds on each of the
8 documents.
\end{enumerate}
%\vspace{-0.5em}
%
If any of (a) or (b) was unsuccessful, the participant could not
continue with the task.
They were allowed to restart from scratch, on a
different unit and therefore on a randomly selected topic, if willing
to do so; the same quality checks were applied again.
If checks (c) or (d) were unsuccessful, the participant received the
message ``\emph{Your job is not accurate enough.
You can revise your work to finish the task}'', and was allowed to use
the back button and revise their previously assigned scores if they
wished (the same two checks (c) and (d) were performed again in such a
case); however, participants were not made aware which documents or
scores were the ``offending'' responses.
Less than 10\% of units resulted in conditions (c) or (d) being
triggered (see details in Table~\ref{tab:descr-stat}, where the number
of eventually successful units with 0, 1, or two or more failed checks
are shown), and as already mentioned the back button was used in less
than 3\% of units.
Finally, there was a syntax check that the numeric scores input by the
participants were in the $(0,+\infty)$ range.
If any of the checks were not successfully completed, no data for that
unit was retained.
Based on these extensive quality checks, we did not carry out 
any further filtering once the data had been collected.

% Local Variables:
% TeX-master: "ME-TOIS.tex"
% End: