### 2016-09-13 ME-TOIS.tex

\documentclass[prodmode,acmtois]{acmsmall}

\usepackage{amsfonts}
\usepackage{amssymb,amsmath}
\usepackage{xcolor}
\usepackage{times}
\usepackage[square,sort,numbers]{natbib}
\setlength{\bibsep}{0.0pt}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{url}
\usepackage{enumerate}
\usepackage[bookmarksnumbered=true,bookmarksopenlevel=4,bookmarksopen=true]{hyperref} 
\usepackage{xcolor}
\definecolor{dark-red}{rgb}{1,0.15,0.15}
\definecolor{dark-blue}{rgb}{0.15,0.15,1}

\hypersetup{
 colorlinks   = true, %Colours links instead of ugly boxes
 urlcolor     = dark-red, %blue Colour for external hyperlinks
 linkcolor    = dark-red, % blueColour of internal links
 citecolor   = dark-blue, %green, %red %Colour of citations
}

\newcommand{\hh}{{\tt H}\xspace}
\newcommand{\rr}{{\tt R}\xspace}
\newcommand{\mm}{{\tt M}\xspace}
\newcommand{\nn}{{\tt N}\xspace}
\newcommand{\nkn}{{\tt N$_k$}\xspace}
\newcommand{\hkh}{{\tt H$_k$}\xspace}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{xcolor}
\usepackage{xargs}  
\usepackage[textsize=small]{todonotes}
%\newcommand{\sm}{\todo[author=\textbf{SM}, color=yellow,inline]}
\newcommandx{\ssm}[2][1=]{\todo[fancyline,author=\textbf{SM},linecolor=yellow,backgroundcolor=yellow,bordercolor=black,#1]{#2}}
%\newcommand{\aht}{\todo[author=\textbf{AHT}, color=green,inline]}
%\newcommand{\fs}{\todo[author=\textbf{FS}, color=cyan,inline]}

%\newcommand{\todo}[1]{\textcolor{red}{[{\bf #1}]}}
\newcommand{\sm}[1]{\textcolor{blue}{[{\bf SM: #1}]}}
\newcommand{\aht}[1]{\textcolor{green}{[{\bf AHT: #1}]}}
\newcommand{\fs}[1]{\textcolor{cyan}{[{\bf FS: #1}]}}
\newcommand{\ehm}[1]{\textcolor{orange}{[{\bf EHM: #1}]}}

%\newcommand{\myparagraph}[1]{\vspace{0.5\baselineskip}\noindent{\textbf{#1}}~}
\newcommand{\defineterm}[1]{\emph{#1}}

%% Avoid floats going to pages / columns by themselves
\renewcommand{\dbltopfraction}{0.9}  % fit big float above 2-col. text
\renewcommand{\textfraction}{0.07}  % allow minimal text w. figs
        %   Parameters for FLOAT pages (not text pages):
\renewcommand{\floatpagefraction}{0.7}  % require fuller float pages
          % N.B.: floatpagefraction MUST be less than topfraction !!
\renewcommand{\dblfloatpagefraction}{0.7} % require fuller float pages

 %\acmVolume{V}
 %\acmNumber{N}
 %\acmArticle{A}
 %\acmYear{YYYY}
 %\acmMonth{0}
\markboth{}{}

\title{On The Benefits of Magnitude Estimation Relevance Assessments
  for Information Retrieval Evaluation}

\title{Crowdsourcing Relevance Magnitudes for Information Retrieval
  Evaluation: On Using Magnitude Estimation to Gather Relevance
  Assessments by Means of Crowdsourcing}

\title{On Crowdsourcing Relevance Magnitudes for Information Retrieval
  Evaluation}

\author{
EDDY MADDALENA
\affil{University of Udine, Italy}
STEFANO MIZZARO
\affil{University of Udine, Italy}
FALK SCHOLER
\affil{RMIT University, Australia}
ANDREW TURPIN
\affil{University of Melbourne, Australia}
}

\begin{document}

\begin{bottomstuff}
Author's addresses: Eddy Maddalena, University of Udine, Udine 33100,
Italy, eddy.maddalena@uniud.it; 
Stefano Mizzaro, University of Udine, Udine 33100, Italy,
mizzaro@uniud.it; 
Falk Scholer. RMIT University, Melbourne, Victoria 3001, Australia,
falk.scholer@rmit.edu.au; 
Andrew Turpin, University of Melbourne, Melbourne, Victoria 3010,
Australia, aturpin@unimelb.edu.au. This paper is a revised and
extended version of \citet{ME-SIGIR15}.
\end{bottomstuff}

\input{0_abstract.tex}

\category{H.3.4}{Information Storage and Retrieval}{Systems and
  software}[performance evaluation] 
\terms{Measurement, performance, experimentation}
\keywords{Magnitude estimation, evaluation, relevance assessments,
  relevance}

\acmformat{...}
\maketitle
% \sm{I think there are good reasons to add ``crowdsourcing'' to the title: the comment analysis is rather related to CS, and it may make sense to add a new CS-related RQ}

\input{1_intro.tex}
\input{2_bg.tex}
\input{3_methods.tex}
\input{4_descriptive_stats.tex}
\input{5_results_RQ1.tex}
\input{6_results_RQ2.tex}
\input{7_results_RQ3.tex}
\input{8_results_RQ4.tex}
%\input{9_comments.tex}
\input{10_conc.tex}
\input{11_appendix.tex}

%Acknowledgments
\begin{acks}
  This work was supported in part by the Australian Research Council
  (\emph{Discovery Project} DP130104007) and in part by a Google Faculty Research Award. 
\end{acks}

%\bibliographystyle{abbrvnat}
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{bib}

%\received{Month Year}{Month Year}{Month Year}

\label{sec:task-steps}
%\input{what-is-new.tex}

\end{document}