### 2016-09-13 a_story.tex

\section{Appendix: NOTES,  TO BE REMOVED}

Outline of story
\begin{itemize}
\item Previous authors have played around with ME and shown it can work.

\item We decided to do it on a larger scale with TREC data to explore the relevance
landscape and look at the effect of judgement variability on gain-based-metrics and system rankings.

\item RQ1: Can ME be used to collect reasonable judgements on a large scale?
        (What does reasonable mean?) Cheap?
        Agreement of pairwise ranks between S, T and ME scores (median and variable) (across and within topics?).
        Across topic variability - does normalisation lead to consistent scores per topic?

        Highlights some "problems" in TREC and Sormunen judgements.

        Explanatory fig: check\_gross\_ranks\_med.pdf (with only
        median-per-doc, rather than all obs per person per doc
        \todo{SM: why not leaving the single colored dots (one dot per
          document, not per judgement, more readable)? They are
          informative  and colors are nice}).
        And across all topics.
        \todo{SM: again, with colored dots?}
        Shows that overall it's sensible, but some disagreements (for
        S, and TREC) and some unjudged that look rel.

        Agreement graph shows that agreement between ME and S
        is about the same as agreement between TREC and S. Also need to
        quantify (maybe put mean/median bars on the graph? And calculate
        an overall mean/median across the cateories, to put in text?

        \todo{Give some examples of why we sometimes see "N" that are
        higher than the median of "H", and vice versa (e.g. we disagree
        with sormunen)}

\item RQ2: Now that we have ``meaningful'' gain values, 
            do system rankings change if you use ME scores as gain vs S categories 
            in NDCG@10 and ERR@10?

\item RQ2.1: using median NME scores for gains

      => tau versus ordinal, and vs TREC

      => ME vs Sorm ~ 0.8; ME vs TREC ~ 0.81, TREC vs Sorm ~ .88

      => Something's going on, look at users?
      (coloured lines graph)

      (=> compare with Kanoulas: "the relative difference between gain
      values is much smaller than the com- monly used ones, giving
      almost equal weights to both rele- vant and highly relevant
      documents" )

\item RQ2.2: prev ignores population varation, try also using resampled user NME scores for gains

      => tau versus ordinal, and vs TREC

\item RQ3: Can ME help us to understand topics better? If you have a population of judges, do they use different gain
      ratios for different topics? AKA do different topics have
      different gain profiles?
      \todo{Histogram of H/N ratios}

      => clearly different for different topics, should not use a single
      gain distribution unlike prev work

      => calls into question the validity of normalising DCG

\item RQ4: Can we use the spread of ME values as a diagnostic for collection/topic/doc quality?
    \begin{itemize}
           \item some topics are ``ambiguous'' across a population of users?
           \item some documents are ``ambiguous'' across a population of users?
    \end{itemize}
    Topic difficulty prediction.

Some other previous RQs follow:

\item RQ2.2: Do system rankings change if you resample ME as gain in NDCG@10 and ERR@10?

\item RQ2.3: What is the range of NDCG@10, ERR@10 you get due to judgement difference?

\item RQ3: Now we know all about users! (ie have numbers/gains), 
           what can we tell about previous categorical gain values?
        Is Sormunen categorisation linear? -> recommend mapping from 4-cats to gain 
              Is it the same across topics?

        Is the TREC categorisation linear? -> recommend mapping from binary to gain (does it matter for binary?)

\item RQ5: But how do I use ME scores for MAP (or any metric requiring a categorical scale)? JUNK!
    \begin{itemize}
        \item RQ3.1 Reduce to binary scale maximising agreement with TREC either per topic, or global.
                    Compute taus. and/or disagreements with TREC.

        \item RQ3.2 Reduce to quad scale maximising agreement with Sor. either per topic, or global.
                    Compute taus. and/or disagreements with Sor.

        \item RQ3.3 Reduce to 2 or 4 scale using k-means with eucllidean.
                    Compute taus. and/or disagreements with Sor.

    \end{itemize}

\end{itemize}