### 2016-09-13 b_stuff.tex

\section{Appendix: OTHER NOTES,  TO BE REMOVED}

********************************************

\todo{Probably remove from here}

********************************************

\subsection{Reliability}
\label{sec:reliability}

Figure~\ref{fig:ME-scores}

\begin{figure}[tb]
  \centering
  \includegraphics[width=\linewidth]{figs/byTopic.pdf}
  \includegraphics[width=\linewidth]{figs/byTopic_docRanges.pdf}
  \caption{ME scores \todo{SM can be removed as it can be inferred by Figure~\ref{fig:ME-distribution-breakdown}}}
  \label{fig:ME-scores}
\end{figure}

\subsubsection{All documents together}
\label{sec:all-docum-togeth}

Figure~\ref{fig:check-ranks}

\begin{figure*}[tbp]
  \centering
%  \includegraphics[width=\textwidth]{figs/check_gross_ranks_all.pdf}
  \caption{All docs together \todo{Commented out this figure for now since it
  is huge and stalls the rendering of the paper pdf!!} \todo{Can be removed and replaced by Figure~\ref{fig:ME-distribution-breakdown}}}
  \label{fig:check-ranks}
\end{figure*}

\subsubsection{Individual topic-document pairs}
\label{sec:indiv-topic-docum}

... Maybe some chart from \verb|8_expertSorted_allNH_only.pdf| and/or \verb|8_expertSorted.pdf|

%% put it in its won file if you prefer
\section{Discussion} 
\label{sec:discussion}

% Local Variables:
% TeX-master: "p.tex"
% End: