### 2016-09-13 letter.tex

\documentclass{letter}
\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{a4wide}

\setlength{\parindent}{0pt}

\makeatletter
\renewcommand*{\opening}[1]{%
%  \ifx\@empty\fromaddress
%    \thispagestyle{firstpage}%
%      {\raggedleft\@date\par}%
%  \else  % home address
%    \thispagestyle{empty}%
%    {\raggedleft\begin{tabular}{l@{}}\ignorespaces
%      \fromaddress \\*[2\parskip]%
%      \@date \end{tabular}\par}%
%  \fi
%  \vspace{2\parskip}%
%  {\raggedright \toname \\ \toaddress \par}%
  \vspace*{-3cm}%
  \begin{minipage}[t]{0.5\textwidth}%
    \toname \\ 
    \toaddress
  \end{minipage}\hfill\@date\par
  \vspace{2\parskip}%
  #1\par\nobreak}
\makeatother

\setlength\textheight{19.5cm}

\begin{document}
\thispagestyle{empty}

\begin{letter}{Editor-in-chief\\ ACM Transactions on Information Systems\\ University of Amsterdam}

\opening{Dear Maarten,}

Thank you for your invitation to submit our SIGIR 2015 paper 
\emph{The Benefits of Magnitude Estimation for Relevance Assessment}
as a fast-tracked TOIS paper.

My co-authors and I have thoroughly revised the paper (in fact, probably a bit too
thoroughly which is why it has taken us so long!), and attach it here
for consideration as a TOIS paper.

In comparison to the SIGIR paper, we have:
\begin{itemize}
    \item Added Research Question 2, addressing the number of crowd
          workers required to make magnitude estimation successful in this context.
    \item Added Section 2.3 discussing crowd sourcing
    \item Added the complete instructions given to workers as an appendix
    \item Added Figures 1 and 2 and accompanying text in Section 4
    \item Reformatted and added explanation to the normalisation method for ME
    \item Added references mentioned by SIGIR reviewers of which we were unaware
    \item Added gsd panel to Figure 5
    \item Completely reworked Section 6.1, including adding Fig 6, recomputing
          Fig 7 differently from SIGIR, Adding Figure 8
    \item Section 6.3 is new, including Figure 9 and Table II
    \item Parts of 7.2 have been reworked, Table III expanded, Figures 
          10 and 11 recomputed using a slightly different (hopefully more robust) method, 
          and Table IV recomputed.
    \item The second half of Section 8 has been reworked, including recomputation of Figure 15.
    \item Section 9 (Conclusions and Future Work) is almost entirely new text.
\end{itemize}

The second page of this letter contains an infographic showing the current TOIS
submission compared with the SIGIR paper computed using \emph{latexdiff}.
Black text is from the SIGIR paper, blue text is added, and red shows
deleted text. Considering both the text and the figures that we have revised, I think
we exceed the required 25\% new material.

I can confirm that all authors are aware of the submission, and agree
to its review by TOIS.

I would suggest Ryen White, Charlie Clarke and Tetsuya Sakai as possible Associate
Editors to handle this submission. No authors have conflict with these AEs.

I hope you find it acceptable for consideration by your reviewers.

Yours Sincerely,

\includegraphics[scale=0.5]{/Users/aturpin/doc/my signature.pdf}

Andrew Turpin\\
Department of Computing and Information Systems\\
University of Melbourne\\
Australia.

\newpage
\thispagestyle{empty}
\includegraphics[scale=0.5]{sigir_in_this_format/diff-nup.jpg}

\end{letter}
\end{document}