### 2016-09-13 what-is-new.tex

\documentclass{article}

 \usepackage{xr}
%\usepackage{xr-hyper,hyperref}
 \usepackage{hyperref}
\externaldocument{ME-TOIS}

% \usepackage{zref-xr}
% \zexternaldocument{ME-TOIS}

\title{What's new}

\begin{document}
\maketitle

\section*{What's new}
%\label{sec:whats-new}

Things that are new w.r.t.\ SIGIR and that we should mention in the
submission letter. 
Let us collect them here, using the power of this xr latex package
that allowa to use Latex references:

\begin{itemize}

\item Overall, the paper is much longer. We have gone from 20 pages
  (once reformatted the SIGIR paper into ACM TOIS style)  to \pageref{sec:task-steps} 
pages, therefore adding more than 40\% (indeed, almost 50\%) new material.
\item The paper now contains a new analysis of several crowdsourcing
  related aspects:
  \begin{itemize}
  \item one new Research Question in Section~\ref{sec-intro} (RQ\ref{item:rq4});
  \item new related work on crowdsourcing (Section~\ref{sec:crowdsourcing});
  \item a new analysis of ME score distributions
    (Section~\ref{sec:score-distribution}, more than 1 novel page);
  \item more details on score normalization and aggregation
    (Section~\ref{sec:score-normalization}, added 1 novel page);
  \item more details on agreement in Section~\ref{sec:judge-agreement}
    (added around 2 novel pages);
  \item the novel analysis on how many workers are needed, in
    Section~\ref{sec:how-many-workers} (more than 1 novel page added).
  \end{itemize}
\item Of the 15 figures in the paper, 6 figures (40\% of them), ie.e., Figures~\ref{fig:ME-raw-scores}, \ref{fig:ME-norm-scores},
  \ref{fig:judgeVariability}, \ref{fig:workersQuality},
  \ref{fig:normaggagain}, and \ref{fig:sampleStandardDeviations}, are
  completely new. Also, 3 figures, i.e., Figures~\ref{fig:agreement},
  \ref{fig:runs-rerank-median3best}, and \ref{fig:runs-rerank-median3best-fake} have been modified.
\item The paper has been restructured in several parts.
\item The references now contain  several new papers, including those previously suggested by referees.
%\item comment analysis, §9
\item We have added an appendix with full details of the crowdsourcing
  instructions, for the sake of replicability. 
\item Everything has been revised.

\end{itemize}
\end{document}